---
title: "Package Test Version 2: New Identification Strategy"
output:
  html_document: default
  html_notebook: default
---


```{r setup, include=FALSE}

require(idealstan)
require(ggplot2)
require(dplyr)
require(tidyr)
library(bayesplot)

knitr::opts_chunk$set(fig.align = 'center',warning = F,message = F)
```


##Introduction

This Rmarkdown file has a series of simulations run on the models offered in `idealstan`. It simulates a large amount of data and then checks to see if the residuals relative to the true parameters sum to zero, as well as for credible interval coverage. In general, the credible intervals aren't going to be exactly where they should be because the constraints used in estimating the models and identifying the signs aren't themselves included in the simulated data. I.e., one of the ideal points actually follows a positive continuous distribution instead of a random normal distribution once it is prefix, which introduces a slight bias in the recovery of the true parameters. This is an artifact of virtually every latent variable model, and is not of great concern as long as the credible intervals reach close to 90 percent coverage. 


First, the standard IRT 2-PL (for this one, the absence discriminations are not a part of the model):

```{r irt_binary}
bin_irt_2pl_sim <- id_sim_gen(num_person=100,num_items=5,ordinal=F,inflate=F)
bin_irt_2pl_est <- id_estimate(idealdata=bin_irt_2pl_sim,
                               model_type=1,init_pathfinder = TRUE,
                               restrict_ind_high = as.character(sort(bin_irt_2pl_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                               restrict_ind_low = as.character(sort(bin_irt_2pl_sim@simul_data$true_person,
                                                                    decreasing=F, 
                                                        index=T)$ix[1]),
                               fixtype='prefix',
                           person_sd = 3,debug_mode = FALSE,
                           nchains=2,ncores=4,niters = 500,warmup=500,
                           compile_optim = F,num_pathfinder_paths = 4)
id_plot_rhats(bin_irt_2pl_est)
```

```{r cov_plot_irt_2pl}
id_sim_coverage(bin_irt_2pl_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

Plot true versus observed values:

```{r irt_2pl_plot}
id_plot_legis(bin_irt_2pl_est,show_true = T)
```


```{r irt_2pl_resid}

id_plot_sims(bin_irt_2pl_est,type='Residuals')

```

```{r irt_2pl_RMSE}

id_plot_sims(bin_irt_2pl_est)

```

```{r irt_2pl_pred}

bin_irt_2pl_est %>% 
  id_post_pred %>% 
  id_plot_ppc(bin_irt_2pl_est,ppc_pred=.,combine_item=FALSE,
              prompt_plot=FALSE)

```


Next, inflated 2-PL IRT (binary):

```{r irt_binary_inflate}
bin_irt_2pl_abs_sim <- id_sim_gen(num_person=20,num_items=100,ordinal=F,inflate=T,
                               absence_diff_mean=0)
bin_irt_2pl_abs_est <- id_estimate(idealdata=bin_irt_2pl_abs_sim,
                               model_type=2,
                               restrict_ind_high = as.character(sort(bin_irt_2pl_abs_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(bin_irt_2pl_abs_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(bin_irt_2pl_abs_est)
```

Plot true versus observed values:

```{r irt_2pl_abs_plot}
id_plot_legis(bin_irt_2pl_abs_est,show_true = T)
```

```{r irt_2pl_abs_cov_plot}
id_sim_coverage(bin_irt_2pl_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r irt_2pl_abs_resid}

id_plot_sims(bin_irt_2pl_abs_est,type='Residuals')

```

```{r irt_2pl_abs_RMSE}

id_plot_sims(bin_irt_2pl_abs_est)

```

```{r irt_2pl_abs_pred}

bin_irt_2pl_abs_est %>% 
  id_post_pred %>% 
  id_plot_ppc(bin_irt_2pl_abs_est,ppc_pred=.,combine_item=FALSE,prompt_plot=FALSE)

```


Now we'll start with the ordinal models. First the uninflated ordinal model:

```{r irt_ordinal}
ord_irt_sim <- id_sim_gen(num_person=20,num_items=100,inflate=F,
                               absence_diff_mean=0,model_type='ordinal_ratingscale')
ord_irt_est <- id_estimate(idealdata=ord_irt_sim,
                               model_type=3,
                               restrict_ind_high = as.character(sort(ord_irt_sim@simul_data$true_person,decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_sim@simul_data$true_person,decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_est)
```

Plot true versus observed values:

```{r irt_ord_plot}
id_plot_legis(ord_irt_est,show_true = T)
```

```{r irt_ord_cov_plot}
id_sim_coverage(ord_irt_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_resid}

id_plot_sims(ord_irt_est,type='Residuals')

```

```{r ord_irt_RMSE}

id_plot_sims(ord_irt_est)

```

```{r ord_irt_pred}

ord_irt_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_est,ppc_pred=.)

```

And we will finish with the inflated ordinal model:

```{r irt_ordinal_abs}
ord_irt_abs_sim <- id_sim_gen(num_person=20,num_items=100,inflate=T,
                               absence_diff_mean=0,model_type='ordinal_ratingscale')
ord_irt_abs_est <- id_estimate(idealdata=ord_irt_abs_sim,
                               model_type=4,
                               restrict_ind_high = as.character(sort(ord_irt_abs_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_abs_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_abs_est)
```

```{r irt_ord_abs_plot}
id_plot_legis(ord_irt_abs_est,show_true = T)
```

```{r ord_irt_abs_cov_plot}
id_sim_coverage(ord_irt_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_abs_resid}

id_plot_sims(ord_irt_abs_est,type='Residuals')

```

```{r ord_irt_abs_RMSE}

id_plot_sims(ord_irt_abs_est)

```

```{r ord_irt_abs_pred}

ord_irt_abs_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_abs_est,ppc_pred=.)

```

We can now try the ordinal graded response version (non-inflated):

```{r irt_ordinal_grm}
ord_irt_grm_sim <- id_sim_gen(num_person=20,num_items=100,inflate=F,
                               absence_diff_mean=0,model_type='ordinal_grm')
ord_irt_grm_est <- id_estimate(idealdata=ord_irt_grm_sim,
                               model_type=5,
                               restrict_ind_high = as.character(sort(ord_irt_grm_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_grm_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_grm_est)
```

```{r ord_irt_grm_cov_plot}
id_sim_coverage(ord_irt_grm_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r irt_ord_grm_plot}
id_plot_legis(ord_irt_grm_est,show_true = T)
```

```{r ord_irt_grm_resid}

id_plot_sims(ord_irt_grm_est,type='Residuals')

```

```{r ord_irt_grm_RMSE}

id_plot_sims(ord_irt_grm_est)

```

```{r ord_irt_grm_pred}

ord_irt_grm_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_grm_est,ppc_pred=.)

```

And now the inflated GRM:

```{r irt_ordinal_grm_abs}
ord_irt_grm_abs_sim <- id_sim_gen(num_person=20,num_items=100,inflate=T,
                               absence_diff_mean=0,model_type='ordinal_grm')
ord_irt_grm_abs_est <- id_estimate(idealdata=ord_irt_grm_abs_sim,
                               model_type=6,
                               restrict_ind_high = as.character(sort(ord_irt_grm_abs_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_grm_abs_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_grm_abs_est)
```

```{r irt_ord_grm_abs_plot}
id_plot_legis(ord_irt_grm_abs_est,show_true = T)
```

```{r ord_irt_grm_abs_cov_plot}
id_sim_coverage(ord_irt_grm_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_grm_abs_resid}

id_plot_sims(ord_irt_grm_abs_est,type='Residuals')

```

```{r ord_irt_grm_abs_RMSE}

id_plot_sims(ord_irt_grm_abs_est)

```

```{r ord_irt_grm_abs_pred}

ord_irt_grm_abs_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_grm_abs_est,ppc_pred=.)

```

# Additional Models V0.3

Models I am adding to this release include the Poisson model:

```{r irt_ordinal_poisson}
ord_irt_poisson_sim <- id_sim_gen(num_person=20,num_items=100,inflate=F,
                               absence_diff_mean=0,model_type='poisson')
ord_irt_poisson_est <- id_estimate(idealdata=ord_irt_poisson_sim,
                               model_type=7,
                               restrict_ind_high = as.character(sort(ord_irt_poisson_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_poisson_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                              fix_high=sort(ord_irt_poisson_sim@simul_data$true_person,
                                                                     decreasing=T)[1],
                              fix_low=sort(ord_irt_poisson_sim@simul_data$true_person,
                                                                   decreasing=F)[1],
                               fixtype='prefix')
id_plot_rhats(ord_irt_poisson_est)
```

```{r irt_poisson_plot}
id_plot_legis(ord_irt_poisson_est,show_true = T)
```

```{r ord_irt_poisson_cov_plot}
id_sim_coverage(ord_irt_poisson_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_poisson_resid}

id_plot_sims(ord_irt_poisson_est,type='Residuals')

```

```{r ord_irt_poisson_RMSE}

id_plot_sims(ord_irt_poisson_est)

```

```{r ord_irt_poisson_pred}

ord_irt_poisson_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_poisson_est,ppc_pred=.)

```



Now Poisson-inflated:

```{r irt_abs_poisson}
ord_irt_poisson_abs_sim <- id_sim_gen(num_person=20,num_items=100,inflate=T,
                               absence_diff_mean=0,model_type='poisson')
ord_irt_poisson_abs_est <- id_estimate(idealdata=ord_irt_poisson_abs_sim,
                               model_type=8,
                               restrict_ind_high = as.character(sort(ord_irt_poisson_abs_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_poisson_abs_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_poisson_abs_est)
```

```{r irt_poisson_abs_plot}
id_plot_legis(ord_irt_poisson_abs_est,show_true = T)
```

```{r ord_irt_poisson_abs_cov_plot}
id_sim_coverage(ord_irt_poisson_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_poisson_abs_resid}

id_plot_sims(ord_irt_poisson_abs_est,type='Residuals')

```

```{r ord_irt_poisson_abs_RMSE}

id_plot_sims(ord_irt_poisson_abs_est)

```



```{r ord_irt_poisson_abs_pred}

ord_irt_poisson_abs_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_poisson_abs_est,ppc_pred=.)

```

```{r ord_irt_poisson_pred_missing}

ord_irt_poisson_abs_est %>% 
  id_post_pred(output='missing') %>% 
  id_plot_ppc(ord_irt_poisson_abs_est,ppc_pred=.)

```

A Normally-distributed outcome with no inflation:

```{r irt_normal}
ord_irt_normal_sim <- id_sim_gen(num_person=20,num_items=100,ordinal=F,inflate=F,
                               absence_diff_mean=0,model_type='normal')
ord_irt_normal_est <- id_estimate(idealdata=ord_irt_normal_sim,
                               model_type=9,
                               restrict_ind_high = as.character(sort(ord_irt_normal_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_normal_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix',compile_optim=F,
                              het_var=FALSE)
id_plot_rhats(ord_irt_normal_est)
```

```{r irt_normal_plot}
id_plot_legis(ord_irt_normal_est,show_true = T)
```

```{r ord_irt_normal_cov_plot}
id_sim_coverage(ord_irt_normal_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_normal_resid}

id_plot_sims(ord_irt_normal_est,type='Residuals')

```

```{r ord_irt_normal_RMSE}

id_plot_sims(ord_irt_normal_est)

```

```{r ord_irt_normal_pred}

ord_irt_normal_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_normal_est,ppc_pred=.)

```

A Normally-distributed outcome with inflation:


```{r irt_normal_abs}
ord_irt_normal_abs_sim <- id_sim_gen(num_person=20,num_items=100,inflate=T,
                               absence_diff_mean=0,model_type='normal')
ord_irt_normal_abs_est <- id_estimate(idealdata=ord_irt_normal_abs_sim,
                               model_type=10,
                               restrict_ind_high = as.character(sort(ord_irt_normal_abs_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_normal_abs_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_normal_abs_est)
```

```{r irt_normal_abs_plot}
id_plot_legis(ord_irt_normal_abs_est,show_true = T)
```

```{r ord_irt_normal_abs_cov_plot}
id_sim_coverage(ord_irt_normal_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_normal_abs_resid}

id_plot_sims(ord_irt_normal_abs_est,type='Residuals')

```

```{r ord_irt_normal_abs_RMSE}

id_plot_sims(ord_irt_normal_abs_est)

```

```{r ord_irt_normal_abs_pred}

ord_irt_normal_abs_est %>% 
  id_post_pred %>% 
  id_plot_ppc(ord_irt_normal_abs_est,ppc_pred=.)

```

```{r ord_irt_normal_abs_miss_pred}

ord_irt_normal_abs_est %>% 
  id_post_pred(output='missing') %>% 
  id_plot_ppc(ord_irt_normal_abs_est,ppc_pred=.)

```

A Lognormal model with no inflation:

```{r irt_lognormal}
ord_irt_lognormal_sim <- id_sim_gen(num_person=20,num_items=100,ordinal=F,inflate=F,
                               absence_diff_mean=0,model_type='lognormal')
ord_irt_lognormal_est <- id_estimate(idealdata=ord_irt_lognormal_sim,
                               model_type=11,
                               restrict_ind_high = as.character(sort(ord_irt_lognormal_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_lognormal_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_lognormal_est)
```

```{r irt_lognormal_plot}
id_plot_legis(ord_irt_lognormal_est,show_true = T)
```


```{r ord_irt_lognormal_cov_plot}
id_sim_coverage(ord_irt_lognormal_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_lognormal_resid}

id_plot_sims(ord_irt_lognormal_est,type='Residuals')

```

```{r ord_irt_lognormal_RMSE}

id_plot_sims(ord_irt_lognormal_est)

```

And last but not least, a lognormal model with inflation:

```{r irt_lognormal_abs}
ord_irt_lognormal_abs_sim <- id_sim_gen(num_person=20,num_items=100,ordinal=F,inflate=F,
                               absence_diff_mean=0,model_type='lognormal')
ord_irt_lognormal_abs_est <- id_estimate(idealdata=ord_irt_lognormal_abs_sim,
                               model_type=12,
                               restrict_ind_high = as.character(sort(ord_irt_lognormal_abs_sim@simul_data$true_person,
                                                                     decreasing=T,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(ord_irt_lognormal_abs_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(ord_irt_lognormal_abs_est)
```

```{r irt_lognormal_abs_plot}
id_plot_legis(ord_irt_lognormal_abs_est,show_true = T)
```

```{r ord_irt_lognormal_abs_cov_plot}
id_sim_coverage(ord_irt_lognormal_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

```{r ord_irt_lognormal_abs_resid}

id_plot_sims(ord_irt_lognormal_abs_est,type='Residuals')

```

```{r ord_irt_lognormal_abs_RMSE}

id_plot_sims(ord_irt_lognormal_abs_est)

```

## Time: Random Walk

We won't test all the possible time auto-correlation settings, only a subset of them. I use an inflated binary model as it is a very common one. 

While these models are identified, it appears that sometimes given the amount of data, they are only weakly identified in the posterior and Rhat values higher than 1.1 are possible (but are not evidence of multi-modality).

First random walks:

```{r irt_binary_time_inflate}
bin_time_irt_2pl_abs_sim <- id_sim_gen(num_person=50,num_items=100,
                                       ordinal=F,inflate=T,
                               absence_diff_mean=0,
                               time_process = 'random',
                               time_points=10)

# as.character(sort(bin_time_irt_2pl_abs_sim@simul_data$true_reg_discrim,
#                                                                   
#                                                                   decreasing=TRUE,
#                                                         index=T)$ix[1])

# as.character(sort(bin_time_irt_2pl_abs_sim@simul_data$true_reg_discrim,
#                                                                    decreasing=F,
#                                                         index=T)$ix[1])

# 5*sort(bin_time_irt_2pl_abs_sim@simul_data$true_reg_discrim,
#                                                                   
#                                                                   decreasing=TRUE)[1]

best_first_time <- 4 

best_second_time <- 500

bin_time_irt_2pl_abs_est <- id_estimate(idealdata=bin_time_irt_2pl_abs_sim,
                               model_type=2,nchains=2,ncores = 6,
                               id_refresh=100,time_var = 10,
                               niters = 500,warmup = 500,const_type = "items",
                          keep_param=list(person_vary=T,person_int=F,item=TRUE,
                                                           extra=T),
                          restrict_sd_high=10,
                          restrict_sd_low=10,
                               vary_ideal_pts = 'random_walk',
                          restrict_var = F,
                          restrict_ind_high = as.character(sort(bin_time_irt_2pl_abs_sim@simul_data$true_reg_discrim,
                                                                     decreasing=TRUE,
                                                        index=TRUE)$ix[1]),
                              restrict_ind_low = as.character(sort(bin_time_irt_2pl_abs_sim@simul_data$true_reg_discrim,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(bin_time_irt_2pl_abs_est) + theme(text=element_text(family=""))
```

```{r irt_2pl_abs_time_cov_plot}
id_sim_coverage(bin_time_irt_2pl_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

Let's do a time-varying ideal point plot with true point overlay:

```{r irt_2pl_abs_time_true_pt}
id_plot_legis_dyn(bin_time_irt_2pl_abs_est,plot_sim = T) + facet_wrap(~person_id,scales="free_y")
```

```{r irt_2pl_time_abs_resid}

id_plot_sims(bin_time_irt_2pl_abs_est,type='Residuals')

```

```{r irt_2pl_time_abs_RMSE}

id_plot_sims(bin_time_irt_2pl_abs_est)

```

## Time: Splines

```{r irt_binary_time_spline_inflate}

bin_time_spline_irt_2pl_abs_sim <- id_sim_gen(num_person=100,num_items=200,inflate=F,
                               absence_diff_mean=0,
                               time_process = 'random',
                               time_points=20,time_sd = .25)

sort_ids_high <- sort(bin_time_spline_irt_2pl_abs_sim@simul_data$true_reg_discrim,
                                                                     decreasing=TRUE,
                                                        index=TRUE)$ix

high_restrict <- c(min(sort_ids_high[1:20]),
                       round(median(sort_ids_high[1:20])),
                   max(sort_ids_high[1:20]))

sort_ids_low <- sort(bin_time_spline_irt_2pl_abs_sim@simul_data$true_reg_discrim,,
                                                                   decreasing=F,
                                                        index=T)$ix

low_restrict <- c(min(sort_ids_low[1:20]),
                       round(median(sort_ids_low[1:20])),
                   max(sort_ids_low[1:20]))

bin_time_spline_irt_2pl_abs_est <- id_estimate(idealdata=bin_time_spline_irt_2pl_abs_sim,
                                           use_groups = F,
                               model_type=1,
                               time_center_cutoff = 50,
                               ncores = parallel::detectCores(),
                               vary_ideal_pts = 'splines',prior_only = FALSE,
                               spline_knots=NULL,spline_degree = 2,
                               const_type = "items",person_sd = 5,
                               diff_reg_sd = 5,
                               discrim_reg_scale = 2,discrim_reg_shape = 2,
                               time_var = .1,restrict_var = F,
                               restrict_ind_high = as.character(high_restrict),
                              restrict_ind_low =as.character(low_restrict),
                               fixtype='prefix',adapt_delta=0.95)
id_plot_rhats(bin_time_spline_irt_2pl_abs_est)
```

```{r irt_2pl_abs_time_spline_cov_plot}
id_sim_coverage(bin_time_spline_irt_2pl_abs_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

Let's do a time-varying ideal point plot with true point overlay:

```{r irt_2pl_abs_time_spline_true_pt}
id_plot_legis_dyn(bin_time_spline_irt_2pl_abs_est,plot_lines = 10) + guides(colour="none") + facet_wrap(~person_id,scales="free_y")
```


```{r irt_2pl_time_ar_spline_resid}

id_plot_sims(bin_time_spline_irt_2pl_abs_est,type='Residuals')

```

```{r irt_2pl_time_spline_abs_RMSE}

id_plot_sims(bin_time_spline_irt_2pl_abs_est)

```



## Time: AR(1)

And autocorrelation:

```{r irt_binary_time_ar_inflate}

bin_time_ar_irt_2pl_abs_sim <- id_sim_gen(num_person=20,num_items=100,inflate=T,
                               absence_diff_mean=0,time_process = 'AR',time_points=20,time_sd = .1)

bin_time_ar_irt_2pl_abs_sim@score_matrix$group_id <- forcats::fct_collapse(bin_time_ar_irt_2pl_abs_sim@score_matrix$person_id,
                                                                           `1`=c("1","2","3"))

bin_time_ar_irt_2pl_abs_est <- id_estimate(idealdata=bin_time_ar_irt_2pl_abs_sim,
                               model_type=2,
                               ncores = parallel::detectCores(),
                               vary_ideal_pts = 'AR1',
                               const_type = "items",
                               restrict_ind_high = as.character(sort(bin_time_ar_irt_2pl_abs_sim@simul_data$true_reg_discrim,
                                                                     decreasing=TRUE,
                                                        index=TRUE)$ix[1]),
                              restrict_ind_low = as.character(sort(bin_time_ar_irt_2pl_abs_sim@simul_data$true_reg_discrim,,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix')
id_plot_rhats(bin_time_ar_irt_2pl_abs_est)
```


## Time: Gaussian Process

The Gaussian process differs in that is a non-parametric approach to time-series modeling. It can fit very flexible curves and also works on irregularly-spaced time series.


```{r time_gp_inflate}

time_gp_sim <- id_sim_gen(num_person=10,num_items=200,inflate=T,
                               absence_diff_mean=0,
                          time_process = 'GP',
                          time_points=20)

# a little harder to constrain, need to figure out minimums

all_low <- apply(time_gp_sim@simul_data$true_person,1,min)

time_gp_est <- id_estimate(idealdata=time_gp_sim,
                               model_type=2,
                           use_vb=F,
                           nchains=3,niters=1000,
                               vary_ideal_pts = 'GP',
                               restrict_ind_high = which(all_low==max(all_low)),
                              restrict_ind_low = which(all_low==min(all_low)),
                               fixtype='prefix')
id_plot_rhats(time_gp_est)
```

```{r time_gp_cov_plot}
id_sim_coverage(time_gp_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

Let's do a time-varying ideal point plot with true point overlay:

```{r time_gp_true_pt}
id_plot_legis_dyn(time_gp_est,plot_sim = T) +facet_wrap(~person_id)
```


```{r time_gp_abs_resid}

id_plot_sims(time_gp_est,type='Residuals')

```

```{r time_gp_abs_RMSE}

id_plot_sims(time_gp_est)

```



## Latent Space Model

Another new addition is the latent space model, which is only a slight modification on the IRT model. 

Non-inflated:

```{r latent_space_noninfl}
latent_noninfl_sim <- id_sim_gen(num_person=10,num_items=125,ordinal=F,
                                       inflate=F,
                                       model_type='binary',
                                       latent_space=T)
latent_noninfl_est <- id_estimate(idealdata=latent_noninfl_sim,
                               model_type=13,
                               restrict_ind_high = as.character(sort(latent_noninfl_sim@simul_data$true_person,
                                                                     decreasing=TRUE,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(latent_noninfl_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix',id_refresh=100,warmup=250,niters=250,nchains=2,ncores=2)
id_plot_rhats(latent_noninfl_est)
```

```{r latent_noninfl_plot}
id_plot_legis(latent_noninfl_est,show_true = T)
```

```{r latent_noninfl_plot_cov}
id_sim_coverage(latent_noninfl_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

Inflated:

```{r latent_space_infl}
# you can't have too few people relative to bills
latent_infl_sim <- id_sim_gen(num_person=20,num_items=125,ordinal=F,
                                       inflate=T,
                                       model_type='binary',
                                       latent_space=T)
latent_infl_est <- id_estimate(idealdata=latent_infl_sim,
                               model_type=14,
                               restrict_ind_high = as.character(sort(latent_infl_sim@simul_data$true_person,
                                                                     decreasing=TRUE,
                                                        index=T)$ix[1]),
                              restrict_ind_low = as.character(sort(latent_infl_sim@simul_data$true_person,
                                                                   decreasing=F,
                                                        index=T)$ix[1]),
                               fixtype='prefix',
                              adapt_delta=0.95,
                              diff_reg_sd = 2,
                             diff_miss_sd=2)
id_plot_rhats(latent_infl_est)
```

```{r latent_infl_plot}
id_plot_legis(latent_infl_est,show_true = T)
```

```{r latent_infl_plot_cov}
id_sim_coverage(latent_infl_est) %>% 
  bind_rows(.id='Parameter') %>% 
  ggplot(aes(y=avg,x=Parameter)) +
  stat_summary(fun.args=list(mult=1.96)) + 
  theme_minimal() 
```

